High resolution, annual maps of field boundaries for smallholder-dominated croplands at national scales

### 2021-10-08 inst/manuscript/manuscript.tex
MS version without EarthArxiv note

\documentclass[11pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{float}  % added 9/3/2021 LDE
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={High resolution, annual maps of the characteristics of smallholder-dominated croplands at national scales},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{5}

@@

\title{High resolution, annual maps of the characteristics of
smallholder-dominated croplands at national scales}
\author{
Lyndon D. Estes
Su Ye
Lei Song
Boka Luo
J. Ronald Eastman
Zhenhua Meng
Qi Zhang
Dennis McRitchie
Stephanie R. Debats
Justus Muhando
Angeline H. Amukoa
Brian W. Kaloo
Jackson Makuru
Ben K. Mbatia
Isaac M. Muasa
Julius Mucha
Adelide M. Mugami
Judith M. Mugami
Francis W. Muinde
Fredrick M. Mwawaza
Jeff Ochieng
Charles J. Oduol
Purent Oduor
Thuo Wanjiku
Joseph G. Wanyoike
Ryan B. Avery
Kelly K. Caylor
}
\date{}
% Allowing for landscape pages
\usepackage{lscape}
\usepackage{graphicx}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}

@@

\begin{document}
%%Edited by @jhollist 09/15/2014
%%Adds title from YAML
\begin{singlespace}
\begin{center}
\huge High resolution, annual maps of the characteristics of
smallholder-dominated croplands at national scales
\end{center}
%%Adds subtitle from YAML
%%Adds Author, correspond email asterisk, and affilnum from YAML
\begin{center}
\large
%% removed space here
Lyndon D. Estes\textsuperscript{*} \textsuperscript{1}, 
%% removed space here
Su Ye \textsuperscript{1,2}, 
%% removed space here
Lei Song \textsuperscript{1}, 
%% removed space here
Boka Luo \textsuperscript{1,3}, 
%% removed space here
J. Ronald Eastman \textsuperscript{1,3}, 
%% removed space here
Zhenhua Meng \textsuperscript{1}, 
%% removed space here
Qi Zhang \textsuperscript{1}, 
%% removed space here
Dennis McRitchie \textsuperscript{4}, 
%% removed space here
Stephanie R. Debats \textsuperscript{4}, 
%% removed space here
Justus Muhando \textsuperscript{5}, 
%% removed space here
Angeline H. Amukoa \textsuperscript{5}, 
%% removed space here
Brian W. Kaloo \textsuperscript{5}, 
%% removed space here
Jackson Makuru \textsuperscript{5}, 
%% removed space here
Ben K. Mbatia \textsuperscript{5}, 
%% removed space here
Isaac M. Muasa \textsuperscript{5}, 
%% removed space here
Julius Mucha \textsuperscript{5}, 
%% removed space here
Adelide M. Mugami \textsuperscript{5}, 
%% removed space here
Judith M. Mugami \textsuperscript{5}, 
%% removed space here
Francis W. Muinde \textsuperscript{5}, 
%% removed space here
Fredrick M. Mwawaza \textsuperscript{5}, 
%% removed space here
Jeff Ochieng \textsuperscript{5}, 
%% removed space here
Charles J. Oduol \textsuperscript{5}, 
%% removed space here
Purent Oduor \textsuperscript{5}, 
%% removed space here
Thuo Wanjiku \textsuperscript{5}, 
%% removed space here
Joseph G. Wanyoike \textsuperscript{5}, 
%% removed space here
Ryan B. Avery \textsuperscript{6}, 
%% removed space here
Kelly K. Caylor \textsuperscript{6,7,8}, 
\end{center}
%%Adds affiliations from YAML
\begin{justify}
% \footnotesize \emph{
% \\*
\footnotesize\textsuperscript{1}Graduate School of Geography, Clark
University, Worcester, MA, USA\\\\*
% \\*
\footnotesize\textsuperscript{2}Department of Natural Resources and the
Environment, University of Connecticut, Storrs, CT, USA\\\\*
% \\*
\footnotesize\textsuperscript{3}Clark Labs, Clark University, Worcester,
MA, USA\\\\*
% \\*
\footnotesize\textsuperscript{4}Independent contributor\\\\*
% \\*
\footnotesize\textsuperscript{5}SpatialCollective, Nairobi, Kenya\\\\*
% \\*
\footnotesize\textsuperscript{6}Department of Geography, University of
California Santa Barbara, Santa Barbara, CA, USA\\\\*
% \\*
\footnotesize\textsuperscript{7}Earth Research Institute, University of
California Santa Barbara, Santa Barbara, CA, USA\\\\*
% \\*
\footnotesize\textsuperscript{8}Bren School of Environmental Science and
Management, University of California Santa Barbara, Santa Barbara, CA,
USA\\\\*
% }
%%Adds corresponding author email(s) from YAML
\newcounter{num}
\setcounter{num}{1}
\\[0.1cm]
\footnotesize \emph{
\ifnum\value{num}=1%
\textsuperscript{*} corresponding author:
\fi
\href{mailto:lestes@clarku.edu}{\nolinkurl{lestes@clarku.edu}}
\stepcounter{num}
}

%\begin{center}
This pre-print not yet undergone peer review. It has been submitted to
\emph{Frontiers in Artificial Intelligence}. This version will be
updated as it is revised, and the final published version will be
accessible through its DOI link.
%\end{center}

\end{justify}
%%Adds date from YAML
\normalsize

\begin{abstract}
Mapping the changing characteristics of Africa's smallholder-dominated
agricultural systems, including the sizes and numbers of fields, is
crucial for understanding food security and a range of other
socioeconomic and environmental concerns. However, accurately mapping
these systems is difficult because of 1) the spatial and temporal
mismatch between satellite sensors and smallholder fields, and 2) the
lack of high-quality labels needed to train and assess machine learning
classifiers. We developed an approach designed to address these two
problems, which we used to map Ghana's annual croplands for the year
2018. To overcome the first problem, we converted daily, high resolution
CubeSat (PlanetScope) imagery into two cloud-free seasonal composites
covering a single agricultural year. To address the second problem, we
created a labelling platform that rigorously assesses and minimizes
label error, and used it to iteratively train a Random Forests
classifier with active learning, which identifies the most informative
training sample based on prediction uncertainty. Minimizing label errors
improved model F1 scores by up to 25\%. Active learning increased F1
scores by an average of 9.1\% between first and last training
iterations, and 2.3\% more than models trained with randomly selected
labels. We used the resulting 3.7 m map of cropland probabilities within
a segmentation algorithm to delineate crop field boundaries. Based on an
independent map reference sample (n=1,207), the cropland probability and
field boundary maps have respective overall accuracies of 88\% and
86.7\%, user's accuracies for the cropland class of 61.2\% and 78.9\%,
and producer's accuracies of 67.3\% and 58.2\%. Using the map reference
sample to calculate an unbiased area estimate from the field boundary
map, we found that cropland covers 17.1\% (15.4-18.9\%) of Ghana. Using
the most accurately digitized labels to calculate and correct for biases
in the segmented field boundaries map, we further estimated the average
size (1.73 ha) and total number (1,662,281) of crop fields in Ghana. Our
results demonstrate an adaptable and transferrable approach for mapping
the characteristics of croplands on an annual basis and over national
extents, with several features that effectively mitigate the errors
inherent in remote sensing of smallholder-dominated agriculture.
\end{abstract}
\end{singlespace}

@@

Amidst all the challenges posed by global change, a particular concern
is how agricultural systems will adapt to meet humanity's growing food
demands, and the impacts that transforming and expanding food systems
will have on societies, economies, and the environment (Searchinger et
al. 2019). A number of efforts are underway to address various aspects
of this challenge, including work on diagnosing and closing yield gaps
(Lobell et al. 2009, e.g. Licker et al. 2010, Mueller et al. 2012),
expanding and commercializing production (Morris and Byerlee 2009), and
to understand (Rulli and D'Odorico 2014, Kehoe et al. 2017, Davis et al.
2020) and mitigate (Estes et al. 2016b) agriculture's ecological
impacts. The success of these efforts depends heavily on data that
accurately describes the location and characteristics of croplands
(Fritz et al. 2015), and, given the rapid pace of agricultural change
(Gibbs et al. 2010, Zeng et al. 2018, Bullock et al. 2021), how these
are changing from one year to the next. Unfortunately, for many regions,
existing cropland datasets are inaccurate, and are usually created as
once-off or infrequently update products. As such, estimates of global
cropland area tend to vary widely, often disagree about where croplands
are located (e.g. Fritz et al. 2011, 2013), and become rapidly outdated.
Errors in these maps can propagate in subsequent analyses that use
cropland data as inputs, resulting in potentially misleading answers
(Estes et al. 2018). Beyond distributions, few data are available on key
cropland characteristics such as field size, an important variable
needed to estimate yield and other key food security variables (Carletto
et al. 2015), and as an indicator of farm size (Levin 2006, Samberg et
al. 2016), a critical component of rural livelihoods given increasing
population densities and longstanding debates about the relationship
between farm size and productivity (Feder 1985, Carletto et al. 2013,
Desiere and Jolliffe 2018).

@@

We generated a catalog of 16232 composite tiles (hereafter simply
``tiles'') for Ghana, consisting of a seasonal pair for each of the 8116
0.05\(^\circ\) tile grid cells covering Ghana. To assess the quality of
the resulting composites, 50 tile grid cells were randomly selected, and
two separate observers graded each corresponding seasonal composite
using a four category that evaluated the degree of 1) residual cloud and
2) cloud shadow, 3) the number of visible scene boundary artifacts, and
4) the proportion of the image with resolution degraded below the 3.7 m
PlanetScope resolution (e.g.~because of between-date image
mis-registrations). Each category was qualitatively ranked from 0-3,
with 0 being the lowest quality, and 3 the highest (see SI for complete
protocol), making the highest possible score 12. We rescaled scores to
fall between 0 and 1.

@@

We ranked the unlabelled primary grid cells from least to most certain,
randomly selected a subset of cells from the top 30\% of the ranking (to
minimize the risk of spatial autocorrelation), and sent these back to
the labelling platform. After these new sites were labelled, they were
added to the starter pool of labels, the model was retrained with the
larger training set, its performance and prediction uncertainty was
reassessed, and a new sample of the most uncertain primary grid cells
was again sent for labelling. This loop repeated until model performance
gains saturated or reached a pre-defined threshold, after which a final
map of cropland probabilities was made for the AOI.

@@

We created a separate labelling instance for each AOI. To create
training and validation labels, labellers (the co-authors of this paper)
logged into the website (built with Flask) for a particular AOI and
navigated to the mapping interface (Figure \ref{fig:labeller}), where
they were presented with a white target box representing a primary grid
cell to label, a set of digitizing tools, and several different sources
of imagery. These included true and false color renderings of the
growing season and dry season PlanetScope composites, and several
virtual globe basemaps. They then used the polygon drawing tool to
digitize the boundaries of all crop fields visible within the
PlanetScope overlays that intersect the target grid cell. For this
project, labellers were instructed to digitize active or recently active
crop fields, avoiding tree crops, and fallow or potentially abandoned
fields (see SI for digitizing rules) To aid with interpretation,
labellers toggled between the PlanetScope renderings and the basemaps to
help form a judgement about what constitutes a field. The labeller
assigned each digitized polygon a class category (e.g.~annual cropland),
saved all completed fields to the database, and were then presented with
the next target to label. If the target grid cell did not contain any
fields, labellers simply pressed save to go to the next cell.

@@

Where \(\theta\) represents the true cover type of a pixel (field or not
field), \emph{D} is the label assigned to that pixel by a labeller, and
\(W_i\) is an individual labeller. P(\(\theta\)\textbar D) is the
probability that the actual cover type is what the labellers who mapped
it says it is, while P(W\(_i\)\textbar D) is the average score (ranging
between 0 and 1) of the accuracy assessment assignments an individual
labeller completed within the AOI, and P(W\(\theta\)\textbar D,
\emph{W}\(_i\)) is the labeller's label for that pixel. This approach
therefore used the average assignment quality score for to weight each
labeller's label for a given pixel (see SI for further details). Each
pixel in the target grid cell was merged using this approach (n =
40000), which helps to minimize individual labellers' errors. We
estimated a confidence measure for each consensus label by calculating
its Bayesian Risk (see SI), which ranges between 0 and 1, with 0
indicating full agreement between labellers for all pixels, and 1
indicating complete disagreement.

@@

We used a combination of
\texttt{GeoTrellis}\footnote{https://github.com/locationtech/geotrellis},
\texttt{rasterio}\footnote{https://rasterio.readthedocs.io/en/latest/},
and \texttt{RasterFrames}\footnote{https://rasterframes.io/} to derive
the features on the fly (which was enabled by converting the composites
to Cloud-optimized Geotiffs\footnote{https://www.cogeo.org/}) and
convert them into Apache Spark DataFrames. The extracted features were
combined with their corresponding training and validation labels and
passed to the machine learning classifier, a \texttt{SparkMLlib}
implementation of Random Forests (Breiman 2001). We trained the model
with a balanced sample and a tree depth of 15 and total tree number of
60, which initial testing showed to provide a reasonable balance between
computational time and performance.

@@

Upon completion of the active learning process, we deployed a five-step
algorithm to create a segmented map of field boundaries. In the first
step, we identified edge features within the imagery. To do this, we
applied the meanshift algorithm (Yizong Cheng 1995) to each dry-season
composite tile, and then passed a Sobel filter over the mean-shifted
green, red, and near-infrared bands, and the corresponding map of
predicted cropland probabilities. We then summed the four resulting edge
images to produce a combined edge image.

In the second step, we used a compact watershed algorithm (Neubert and
Protzel 2014) to segment the edge image, specifying a high number of
segments (6,400) segments per tile, so that the mean segment size
(\textless0.5 ha) was finer than the expected mean field size
(\textgreater1 ha).

@@

We used the sample to calculate the overall accuracy for each map, the
class-wise User's and Producer's accuracy, and the 95\% confidence
intervals for each accuracy measure (Olofsson et al. 2013, 2014, Stehman
and Foody 2019). We calculated these measures across the entire country,
as well as several different zones, to evaluate regional difference in
accuracy. We defined two sets of zonations (Figure S4), each containing
four zones, the first created by grouping 1) the three northern AOIs
(1-3), 2) the six central AOIs (4-9), 3) the four southwestern AOIs (10,
11, 13, 14, 16), and 4) the two southeastern zones (13, 15). This
grouping differs from the three clusters used to collect initial model
training samples, as we designed these to divide the country more
finely, and to isolate the less forested southeastern third of Ghana
from the more forest northwest. The second zonation was developed by
grouping the country's eight agro-ecological zones into four broader
clusters (Figure S4B). We applied this zonation only to the per-pixel
classification, to better understand patterns of error in the model.

To assess how effectively the segmentations captured field
characteristics, we compared the size class distributions of the
segmented field boundaries against the field boundaries digitized by the
labellers' over the 100 validation sites in each AOI. We chose this
approach because of existing uncertainties in polygon-based accuracy
assessment methods (Ye et al. 2018), and because the map's ability to
represent field sizes was of greatest interest. To undertake this
comparison, we selected the polygons from the most accurate labeller to
digitize the 100 validation grids in each AOI, and calculated the
average area and number of polygons in each cell. We then calculated the
same statistics from the segmented boundaries that intersected each
validation grid, and compared the two sets of statistics.

@@

The assessment of image composites found that their quality in both
seasons was highest in the northern half of the country and lowest in
the southwest, (Figure \ref{fig:imqual}A), where the substantially
greater cloud cover resulted in a much lower density of available
PlanetScope imagery for each time period (Figure S5). The average
quality score of growing season composites was 0.88, with 70 percent
having scores \(\geq\) 0.85 (out of 1; Figure \ref{fig:imqual}B), while
the mean score of dry season composites was 0.92 (74 percent \(\geq\)
0.85).

@@

To make the initial maps of cropland probabilities, the active learning
process ran for 3 iterations in 12 of 16 AOIs, varying from as little as
1 to as many as 4 iterations across the other 4 AOIs, with the number of
iterations varying according to the performance of the starter models
(i.e.~AOIs with higher starting performance stopped after fewer
iterations, see SI). Each AOI's model was trained by 300-500 randomly
selected labels (Figure S6A), plus an additional 600 - 900 (typically
800) within each the AOI that were selected by active learning. Actively
selected labels showed distinctive patterns in several AOIs (Figure
S6B), such as concentrating along ecotones or the boundaries of
agro-ecological zones. A total of 6,299 training and 1,600 validation
labels were collected by 20 labellers to develop and assess model
performance (Figure S7).

>>>>

\documentclass[11pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{float}  % added 9/3/2021 LDE
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={High resolution, annual maps of field boundaries for smallholder-dominated croplands at national scales},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{5}

@@

\title{High resolution, annual maps of field boundaries for
smallholder-dominated croplands at national scales}
\author{
Lyndon D. Estes
Su Ye
Lei Song
Boka Luo
J. Ronald Eastman
Zhenhua Meng
Qi Zhang
Dennis McRitchie
Stephanie R. Debats
Justus Muhando
Angeline H. Amukoa
Brian W. Kaloo
Jackson Makuru
Ben K. Mbatia
Isaac M. Muasa
Julius Mucha
Adelide M. Mugami
Judith M. Mugami
Francis W. Muinde
Fredrick M. Mwawaza
Jeff Ochieng
Charles J. Oduol
Purent Oduor
Thuo Wanjiku
Joseph G. Wanyoike
Ryan B. Avery
Kelly K. Caylor
}
\date{}
% Allowing for landscape pages
\usepackage{lscape}
\usepackage{graphicx}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}

@@

\begin{document}
%%Edited by @jhollist 09/15/2014
%%Adds title from YAML
\begin{singlespace}
\begin{center}
\huge High resolution, annual maps of field boundaries for
smallholder-dominated croplands at national scales
\end{center}
%%Adds subtitle from YAML
%%Adds Author, correspond email asterisk, and affilnum from YAML
\begin{center}
\large
%% removed space here
Lyndon D. Estes\textsuperscript{*} \textsuperscript{1}, 
%% removed space here
Su Ye \textsuperscript{1,2}, 
%% removed space here
Lei Song \textsuperscript{1}, 
%% removed space here
Boka Luo \textsuperscript{1,3}, 
%% removed space here
J. Ronald Eastman \textsuperscript{1,3}, 
%% removed space here
Zhenhua Meng \textsuperscript{1}, 
%% removed space here
Qi Zhang \textsuperscript{1}, 
%% removed space here
Dennis McRitchie \textsuperscript{4}, 
%% removed space here
Stephanie R. Debats \textsuperscript{4}, 
%% removed space here
Justus Muhando \textsuperscript{5}, 
%% removed space here
Angeline H. Amukoa \textsuperscript{5}, 
%% removed space here
Brian W. Kaloo \textsuperscript{5}, 
%% removed space here
Jackson Makuru \textsuperscript{5}, 
%% removed space here
Ben K. Mbatia \textsuperscript{5}, 
%% removed space here
Isaac M. Muasa \textsuperscript{5}, 
%% removed space here
Julius Mucha \textsuperscript{5}, 
%% removed space here
Adelide M. Mugami \textsuperscript{5}, 
%% removed space here
Judith M. Mugami \textsuperscript{5}, 
%% removed space here
Francis W. Muinde \textsuperscript{5}, 
%% removed space here
Fredrick M. Mwawaza \textsuperscript{5}, 
%% removed space here
Jeff Ochieng \textsuperscript{5}, 
%% removed space here
Charles J. Oduol \textsuperscript{5}, 
%% removed space here
Purent Oduor \textsuperscript{5}, 
%% removed space here
Thuo Wanjiku \textsuperscript{5}, 
%% removed space here
Joseph G. Wanyoike \textsuperscript{5}, 
%% removed space here
Ryan B. Avery \textsuperscript{6}, 
%% removed space here
Kelly K. Caylor \textsuperscript{6,7,8}, 
\end{center}
%%Adds affiliations from YAML
\begin{justify}
% \footnotesize \emph{
% \\*
\footnotesize\textsuperscript{1}Graduate School of Geography, Clark
University, Worcester, MA, USA\\\\*
% \\*
\footnotesize\textsuperscript{2}Department of Natural Resources and the
Environment, University of Connecticut, Storrs, CT, USA\\\\*
% \\*
\footnotesize\textsuperscript{3}Clark Labs, Clark University, Worcester,
MA, USA\\\\*
% \\*
\footnotesize\textsuperscript{4}Independent contributor\\\\*
% \\*
\footnotesize\textsuperscript{5}SpatialCollective, Nairobi, Kenya\\\\*
% \\*
\footnotesize\textsuperscript{6}Department of Geography, University of
California Santa Barbara, Santa Barbara, CA, USA\\\\*
% \\*
\footnotesize\textsuperscript{7}Earth Research Institute, University of
California Santa Barbara, Santa Barbara, CA, USA\\\\*
% \\*
\footnotesize\textsuperscript{8}Bren School of Environmental Science and
Management, University of California Santa Barbara, Santa Barbara, CA,
USA\\\\*
% }
%%Adds corresponding author email(s) from YAML
\newcounter{num}
\setcounter{num}{1}
\\[0.1cm]
\footnotesize \emph{
\ifnum\value{num}=1%
\textsuperscript{*} corresponding author:
\fi
\href{mailto:lestes@clarku.edu}{\nolinkurl{lestes@clarku.edu}}
\stepcounter{num}
}
\end{justify}
%%Adds date from YAML
\normalsize

\begin{abstract}
Mapping the characteristics of Africa's smallholder-dominated croplands,
including the sizes and numbers of fields, can provide critical insights
into food security and a range of other socioeconomic and environmental
concerns. However, accurately mapping these systems is difficult because
there is 1) a spatial and temporal mismatch between satellite sensors
and smallholder fields, and 2) a lack of high-quality labels needed to
train and assess machine learning classifiers. We developed an approach
designed to address these two problems, and used it to map Ghana's
croplands. To overcome the spatio-temporal mismatch, we converted daily,
high resolution imagery into two cloud-free composites (the primary
growing season and subsequent dry season) covering the 2018 agricultural
year, providing a seasonal contrast that helps to improve classification
accuracy. To address the problem of label availability, we created a
platform that rigorously assesses and minimizes label error, and used it
to iteratively train a Random Forests classifier with active learning,
which identifies the most informative training sample based on
prediction uncertainty. Minimizing label errors improved model F1 scores
by up to 25\%. Active learning increased F1 scores by an average of
9.1\% between first and last training iterations, and 2.3\% more than
models trained with randomly selected labels. We used the resulting 3.7
m map of cropland probabilities within a segmentation algorithm to
delineate crop field boundaries. Using an independent map reference
sample (n=1,207), we found that the cropland probability and field
boundary maps had respective overall accuracies of 88\% and 86.7\%,
user's accuracies for the cropland class of 61.2\% and 78.9\%, and
producer's accuracies of 67.3\% and 58.2\%. An unbiased area estimate
calculated from the map reference sample indicates that cropland covers
17.1\% (15.4-18.9\%) of Ghana. Using the most accurate validation labels
to correct for biases in the segmented field boundaries map, we
estimated that the average size and total number of field in Ghana are
1.73 ha and 1,662,281, respectively. Our results demonstrate an
adaptable and transferable approach for developing annual, country-scale
maps of crop field boundaries, with several features that effectively
mitigate the errors inherent in remote sensing of smallholder-dominated
agriculture.
\end{abstract}
\end{singlespace}

@@

Amidst all the challenges posed by global change, a particular concern
is how agricultural systems will adapt to meet humanity's growing food
demands, and the impacts that transforming and expanding food systems
will have on societies, economies, and the environment (Searchinger et
al. 2019). A number of efforts are underway to address various aspects
of this challenge, including work on diagnosing and closing yield gaps
(Lobell et al. 2009, e.g. Licker et al. 2010, Mueller et al. 2012),
expanding and commercializing production (Morris and Byerlee 2009), and
to understand (Rulli and D'Odorico 2014, Kehoe et al. 2017, Davis et al.
2020) and mitigate (Estes et al. 2016b) agriculture's ecological
impacts. The success of these efforts depends heavily on data that
accurately describes the location and characteristics of croplands
(Fritz et al. 2015), and, given the rapid pace of agricultural change
(Gibbs et al. 2010, Zeng et al. 2018, Bullock et al. 2021), how these
are changing from one year to the next. Unfortunately, for many regions,
existing cropland datasets are inaccurate, and are usually created as
once-off or infrequently updated products. As such, estimates of global
cropland area tend to vary widely, often disagree about where croplands
are located (e.g. Fritz et al. 2011, 2013), and become rapidly outdated.
Errors in these maps can propagate in subsequent analyses that use
cropland data as inputs, resulting in potentially misleading answers
(Estes et al. 2018). Beyond distributions, few data are available on key
cropland characteristics such as field size, an important variable
needed to estimate yield and other key food security variables (Carletto
et al. 2015), and as an indicator of farm size (Levin 2006, Samberg et
al. 2016), a critical component of rural livelihoods given increasing
population densities and longstanding debates about the relationship
between farm size and productivity (Feder 1985, Carletto et al. 2013,
Desiere and Jolliffe 2018).

@@

We generated a catalog of 16232 composite tiles (hereafter simply
``tiles'') for Ghana, consisting of a seasonal pair for each of the 8116
0.05\(^\circ\) tile grid cells covering Ghana. To assess the quality of
the resulting composites, 50 tile grid cells were randomly selected, and
two separate observers graded each corresponding seasonal composite
using four categories that evaluated the degree of 1) residual cloud and
2) cloud shadow, 3) the number of visible scene boundary artifacts, and
4) the proportion of the image with resolution degraded below the 3.7 m
PlanetScope resolution (e.g.~because of between-date image
mis-registrations). Each category was qualitatively ranked from 0-3,
with 0 being the lowest quality, and 3 the highest (see SI for complete
protocol), making the highest possible score 12. We rescaled scores to
fall between 0 and 1.

@@

We ranked the unlabelled primary grid cells from least to most certain,
randomly selected a subset of cells from the top 30\% of the ranking (to
minimize the risk of spatial autocorrelation), and sent these back to
the labelling platform. After these new sites were labelled, they were
added to the starter pool of labels, the model was retrained with the
larger training set, its performance and prediction uncertainty was
reassessed, and a new sample of the most uncertain primary grid cells
was again sent for labelling. This loop was typically repeated for 3
iterations, after which a final map of cropland probabilities was made.

@@

We created a separate labelling instance for each AOI. To create
training and validation labels, labellers (the co-authors of this paper)
logged into the website (built with Flask) for a particular AOI and
navigated to the mapping interface (Figure \ref{fig:labeller}), where
they were presented with a white target box representing a primary grid
cell to label, a set of digitizing tools, and several different sources
of imagery. These included true and false color renderings of the
growing season and dry season PlanetScope composites, and several
virtual globe basemaps. They then used the polygon drawing tool to
digitize the boundaries of all crop fields visible within the
PlanetScope overlays that intersect the target grid cell. For this
project, labellers were instructed to digitize active or recently active
crop fields, avoiding tree crops, and fallow or potentially abandoned
fields (see SI for digitizing rules). To aid with interpretation,
labellers toggled between the PlanetScope renderings and the basemaps to
help form a judgement about what constitutes a field. The labeller
assigned each digitized polygon a class category (e.g.~annual cropland),
saved all completed fields to the database, and were then presented with
the next target to label. If the target grid cell did not contain any
fields, labellers simply pressed save to go to the next cell.

@@

Where \(\theta\) represents the true cover type of a pixel (field or not
field), \emph{D} is the label assigned to that pixel by a labeller, and
\(W_i\) is an individual labeller. P(\(\theta\)\textbar D) is the
probability that the actual cover type is what the labellers who mapped
it says it is, while P(W\(_i\)\textbar D) is the average score (ranging
between 0 and 1) of the accuracy assessment assignments an individual
labeller completed within the AOI, and P(W\(\theta\)\textbar D,
\emph{W}\(_i\)) is the labeller's label for that pixel. This approach
therefore used the average assignment quality score to weight each
labeller's label for a given pixel (see SI for further details). Each
pixel in the target grid cell was merged using this approach (n =
40000), which helps to minimize individual labellers' errors. We
estimated a confidence measure for each consensus label by calculating
its Bayesian Risk (see SI), which ranges between 0 and 1, with 0
indicating full agreement between labellers for all pixels, and 1
indicating complete disagreement.

@@

We used a combination of
\texttt{GeoTrellis}\footnote{https://github.com/locationtech/geotrellis},
\texttt{rasterio}\footnote{https://rasterio.readthedocs.io/en/latest/},
and \texttt{RasterFrames}\footnote{https://rasterframes.io/} to derive
the features on the fly (which was enabled by converting the composites
to Cloud-optimized Geotiffs\footnote{https://www.cogeo.org/}) and
convert them into Apache Spark DataFrames.

The extracted features were combined with their corresponding training
and validation labels and passed to the machine learning classifier, a
\texttt{SparkMLlib} implementation of Random Forests (Breiman 2001). We
trained the model with a balanced sample and a tree depth of 15 and
total tree number of 60. Initial testing showed that model performance
saturated with increasing values of these parameters (cluster failures
occurred when tree depths and numbers were simultaneously \(\geq16\) and
\(\geq50\), respectively), and that model stability was satisfactory
with these settings, as there was \(\leq\) 0.01 difference in accuracy
for separate models trained on the same labels.

@@

Upon completion of the active learning process, we deployed a five-step
algorithm to create a segmented map of field boundaries (see Figure S3
for illustration of the steps). In the first step, we identified edge
features within the imagery. To do this, we applied the meanshift
algorithm (Yizong Cheng 1995) to each dry-season composite tile, and
then passed a Sobel filter over the mean-shifted green, red, and
near-infrared bands, and the corresponding map of predicted cropland
probabilities. We then summed the four resulting edge images to produce
a combined edge image.

In the second step, we used a compact watershed algorithm (Neubert and
Protzel 2014) to segment the edge image, specifying a high number of
segments (6,400) per tile, so that the mean segment size (\textless0.5
ha) was finer than the expected mean field size (\textgreater1 ha).

@@

We used the sample to calculate the overall accuracy for each map, the
class-wise User's and Producer's accuracy, and the 95\% confidence
intervals for each accuracy measure (Olofsson et al. 2013, Olofsson et
al. 2014, Stehman and Foody 2019). We calculated these measures across
the entire country, as well as several different zones, to evaluate
regional difference in accuracy. We defined two sets of zonations
(Figure S5), each containing four zones, the first created by grouping
1) the three northern AOIs (1-3), 2) the six central AOIs (4-9), 3) the
four southwestern AOIs (10, 11, 13, 14, 16), and 4) the two southeastern
zones (13, 15). This grouping differs from the three clusters used to
collect initial model training samples, as we designed these to divide
the country more finely, and to isolate the less forested southeastern
third of Ghana from the more forest northwest. The second zonation was
developed by grouping the country's eight agro-ecological zones into
four broader clusters (Figure S5B). We applied this zonation only to the
per-pixel classification, to better understand patterns of error in the
model.

To assess how effectively the segmentations captured field
characteristics, we compared the size class distributions of the
segmented field boundaries against those calculated from the field
boundaries digitized by the labellers within the 100 validation sites
from each AOI. We chose this approach because of existing uncertainties
in polygon-based accuracy assessment methods (Ye et al. 2018), and
because the map's ability to represent field sizes was of greatest
interest. To undertake this comparison, we selected the polygons from
the most accurate labeller to digitize the 100 validation grids in each
AOI, and calculated the average area and number of polygons in each
cell. We then calculated the same statistics from the segmented
boundaries that intersected each validation grid, and compared the two
sets of statistics.

@@

The assessment of image composites found that their quality in both
seasons was highest in the northern half of the country and lowest in
the southwest, (Figure \ref{fig:imqual}A), where the substantially
greater cloud cover resulted in a much lower density of available
PlanetScope imagery for each time period (Figure S6). The average
quality score of growing season composites was 0.88, with 70 percent
having scores \(\geq\) 0.85 (out of 1; Figure \ref{fig:imqual}B), while
the mean score of dry season composites was 0.92 (74 percent \(\geq\)
0.85).

@@

To make the initial maps of cropland probabilities, the active learning
process ran for 3 iterations in 12 of 16 AOIs, varying from as little as
1 to as many as 4 iterations across the other 4 AOIs, with the number of
iterations varying according to the performance of the starter models
(i.e.~AOIs with higher starting performance stopped after fewer
iterations, see SI). Each AOI's model was trained by 300-500 randomly
selected labels (Figure S7A), plus an additional 600 - 900 (typically
800) labels within each AOI that were selected by active learning.
Actively selected labels showed distinctive patterns in several AOIs
(Figure S7B), such as concentrating along ecotones or the boundaries of
agro-ecological zones. A total of 6,299 training and 1,600 validation
labels were collected by 20 labellers to develop and assess model
performance (Figure S8).
